Eclipse Che for Fair and Easy Education

### 2019-05-04 paper.tex
Re-Implemented Citing

Figure \ref{fig:digdocker} shows the steadily rising interest in the most popular cloud technologies. 
Docker is a container-engine enabling OS-Virtualization, Kubernetes is a Container-Platform manager and Openshift is a software-suite based on Kubernetes. 
These technologies will be summarized in \ref{subsec:Req}.  
As cloud-technology such technologies are referred which achieve five attributes: 
On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service [15]. 
Cloud technology usually splits into two components:  Infrastructure  and  Services,  where  docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
Most  of  the  attributes  are (pseudo-) achievable by the infrastructure alone – such as restarting containers. 
Software running as a service can be put into two categories: 
“Cloud-Ready” – meaning that it can be run in common cloud environments, such as Kubernetes. 
The term cloudready  does  usually  not  include  live-scalability. Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a dockercontainer. 
Lately the term “Cloud-native” came up [17], referring to software which is built from the very first steps to run inside the cloud. This kind of software is usually running inside a docker-container and both resilient and elasticby design [8].  
Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
Elasticy is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
Cloud-ready Software usually achieves only a certain degree of resilience.   
Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities [16]. 
The applications must also be able to use the gained resources in a productive and healthy way.  
Modern  cloud-platforms  can  perform  pseudo-scalabilty for cloud-ready-services by starting multiple instances of the same service. 
The cloud-platform performs load-balancing between these instances. 
This naive approach can work out fine, but for example starting two services working on the same database will not yield any real scalability. 
Another common problem arises when two services are started on the same machine and need the same port, just to name two problems with cloud-ready  software in a cloud environment.

As Cloud is a central element of modern IT-development, it is strange to see that it is not yet widely teached across 
universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud-native development is sealed behind corporate doors and conference workshops.
Meanwhile,  the  annual  Stackoverflow  developersurvey [22] showed, that DevOps-engineers are not only 
amongst the most wanted and most paid jobs but also tend to be the happiest participants.  
Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career. 
So why are there no courses? There are many reasons, but the most prominent is the complexity of the topic: 
To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
infrastructure  and  networking. 
Additionally, the developer needs to use specific tools, know best practices, tests and work in a team.
All these factors make a high stake for entrance, also  represented in the developersurvey [22] which states that most DevOps engineers have a decade of experience, usually in operations topics.
Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
Che is unlike a normal eclipse distribution: 
It is hosted on a server and developers get access with their browser. 
Instead of installing dependencies, compilers etc. locally, a common workspace is set up, which is shared amongst the programmers. 
The software build is run inside a docker-container, making it cloud-native by default. 
With Version 7 upcoming and prominent support from RedHat, Eclipse Che is not a prototype anymore. 
The latest distributions of RedHats OpenShift are shipped with Che in default. 
Therefore, it is worth looking at this possible game-changer. 
With the first part of this paper, the structure and ideas of Che are explained in more detail, the second part of this pa-
per covers general arguments about this approach and gives a detailed overview of reachable benefits for education.

@@

Openshift\footnote{https://www.openshift.com/} provides a software-suite around Kubernetes with the goal to automate cloud-native development and delivery.
Notable  additions  include  Jenkins,  Gitea,  Sonarqube and lately Che. 
Regarding Che there are no notable changes in handling, as Che is simply run on the build in Kubernetes.
When run on Kubernetes or Docker Che will require two additional containers for authentification with KeyCloack. 
For Openshift the standard-openshift authentification is used.  
Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations aswell as for developers working on the Che-Code. 
When run (locally) on docker, the plugin CheDir enables a portable workspace, 
which can be used on any docker based Che.
\subsection{Technology and Workflow}
\label{subsec:Tec}
A normal IDE is software on your computer which makes programming easier than doing it in a plain text-editor and commandline. 
Common tasks for an IDE are dependency management, debugging, refactoring and autocompletion.
It is important that an IDE is not required to develop – you can write .java-files in Vim, compile on the console and run a .Jar manually. 
With that in mind, we can see that the IDE is just a tool to make things (a lot) more comfortable – but you also have requirements such as the JDK, the .Jars you need in your classpath etc. as they are only utilized by your IDE.
You can consider this stack as your workspace as it contains everything required to work, with the IDE making the workspace comfortable to use.
Eclipse Che does not work like a normal IDE. 
Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source:\url{https://en.wikipedia.org/wiki/Eclipse_Che#/media/File:Eclipse_Che_-_Workflow.PNG}}.

@@

The university of Arizona showed [11], that about 50\% of software inspected was not able to build. 

@@

Experienced programmers are usually able to overcome these issues, given enough time. This continues to be an issue [14], and without a different approach probably always will be. 
This problem is solved, if the software is built inside a docker-container and follows the principles of self-containment and image-immutability [17].

@@

Che therefore embraces the core-ideas of DevOps [19][20] and agile software development [18].

@@

This is one of the primary goals of the DevOps approach [20]. 
The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of microservices [21], while being agile and reproducible.

@@

Clark et al. [14] describe this rather as a change in the mindset, than in actual technology:

@@

Another problem with Che is the uncommon debugging. 
As the code is running in a remote-runtime, the common breakpoint  and  valuecheck
is  missing.  
Instead,  one  either must be familiar with the languages commandline-debugging or write good logs and metrics. 
While this will slow down development, it also offers some benefits: 
The debugging and error-analysis inside the container is the same as it would be in the productive environment. 
Having meaningful logs with different loglevels and collecting valuable metrics is 
in general a good attribute for software in any circumstance. 
It just means additional effort and know-how.  
The last problem is the vulnerability to malevolent users. 
Someone who already is in possession of a valid account can drain the shared resources of a full cluster, therefore blocking  any  other  user  from  work.
This  can  also  be achieved by human error. 
Dependent on the setup, the attacker could also auto-deploy hidden spyware into a pro-
duction environment. 
While this is also a problem for normal developer teams, the fast pace of deployments with Che requires additional care. 
A rich CI/CD environment, such as a well-maintained Jenkins may lead to less-detailed code-reviews, as both unit- and end-to-end-tests succeed.  
It is therefore important to not only restrict access to project resources, but also the double-check and guard the releases.  
Some security risks can also emerge from docker-images as well [9] but are out of scope for this paper.

@@

Even complicated topics, like microservices and IoT-applications can easily be adaptet inside Che [21] and performed as for example as group-projects with a common goal. 

@@

With both the product and the code inside the workspace easily accessible, the professor can grant faster and better feedback for the students, which can have massive impact on learning [1][5].

@@

This immediate feedback is a big motivator and has a positive impact on the learning experience [2]. 
With a clear reason on the usefulness of the task, the visible progress and the direct impact of personal effort should lead to intrinsic motivation, which studies show to have a strong effect on education [23]. 

@@

Other studies have shown that only automated and unified testing is not beneficial to education in general [3]. 

@@

The topic of socio-economic factors is beyond the scope of this work, but the impact has been analyzed by the PISA Studies [7], which state that socio-economicly handicapped students fail three times more likely to achieve a basic level of performance.

@@

Germany scored on the impact of socioeconomic-factors on education better than it did in 2006 but is worse than global average [6].

@@

Making  assignments  mandatory, providing extrinsic motivation [23] has usually a negative impact on the subjective experience.

@@

Long waiting times disconnect the action from the result being not beneficial for the learning process [1][23]. 

@@

With studies showing basics of good learning experiences [2][3] and the impact of motivation [1][23], it is mandatory for institutions to provide the best possible environment. 

>>>>

Figure \ref{fig:digdocker} shows the steadily rising interest in the most popular cloud technologies. 
Docker is a container-engine enabling OS-Virtualization, Kubernetes is a Container-Platform manager and Openshift is a software-suite based on Kubernetes. 
These technologies will be summarized in \ref{subsec:Req}.  
As cloud-technology such technologies are referred which achieve five attributes: 
On-demand self-service, broad network-access, resource pooling, rapid elasticity and measured service \cite{ME11}. 
Cloud technology usually splits into two components:  Infrastructure  and  Services,  where  docker and Kubernetes are considered infrastructure and provide a  platform  to  host  services.  
Most  of  the  attributes  are (pseudo-) achievable by the infrastructure alone – such as restarting containers. 
Software running as a service can be put into two categories: 
“Cloud-Ready” – meaning that it can be run in common cloud environments, such as Kubernetes. 
The term cloudready  does  usually  not  include  live-scalability. Cloud-readieness is mostly achieved by removing dependencies and either running on bare-linux or being ported into a dockercontainer. 
Lately the term “Cloud-native” came up \cite{BI17}, referring to software which is built from the very first steps to run inside the cloud. This kind of software is usually running inside a docker-container and both resilient and elasticby design \cite{TO17}.  
Resilience refers to the ability of handling failure, both from external sources and restarting/configuring themselves.  
Elasticy is the ability to request more resources on demand, performing more computations when required, but also to free the resources if not needed. 
Cloud-ready Software usually achieves only a certain degree of resilience.   
Cloud-Native applications are aware of their context and collect meaningful metrics for the platform that they cannot reach other services or need more capacities \cite{MIN18}. 
The applications must also be able to use the gained resources in a productive and healthy way.  
Modern  cloud-platforms  can  perform  pseudo-scalabilty for cloud-ready-services by starting multiple instances of the same service. 
The cloud-platform performs load-balancing between these instances. 
This naive approach can work out fine, but for example starting two services working on the same database will not yield any real scalability. 
Another common problem arises when two services are started on the same machine and need the same port, just to name two problems with cloud-ready  software in a cloud environment.

As Cloud is a central element of modern IT-development, it is strange to see that it is not yet widely teached across 
universities in Germany \footnote{https://www.hochschulkompass.de/studium/studiengangsuche lists currently only 10 german universities with courses on cloud computing, about 30 courses on virtualization }.
There are courses to teach either classic virtualization or docker-basics, but the knowledge of full-stack cloudplatforms and cloud-native development is sealed behind corporate doors and conference workshops.
Meanwhile,  the  annual  Stackoverflow  developersurvey \cite{SO19} showed, that DevOps-engineers are not only 
amongst the most wanted and most paid jobs but also tend to be the happiest participants.  
Every student either wants to be happy or to be rich, making Cloud an interesting topic for their career. 
So why are there no courses? There are many reasons, but the most prominent is the complexity of the topic: 
To build a real cloud native application, the developer must understand virtualization, system-administration, development, 
infrastructure  and  networking. 
Additionally, the developer needs to use specific tools, know best practices, tests and work in a team.
All these factors make a high stake for entrance, also  represented in the developersurvey \cite{SO19} which states that most DevOps engineers have a decade of experience, usually in operations topics.
Lucky for us, a new open-source software, Eclipse Che, is on the rise to enable everyone for cloud-native development.
Che is unlike a normal eclipse distribution: 
It is hosted on a server and developers get access with their browser. 
Instead of installing dependencies, compilers etc. locally, a common workspace is set up, which is shared amongst the programmers. 
The software build is run inside a docker-container, making it cloud-native by default. 
With Version 7 upcoming and prominent support from RedHat, Eclipse Che is not a prototype anymore. 
The latest distributions of RedHats OpenShift are shipped with Che in default. 
Therefore, it is worth looking at this possible game-changer. 
With the first part of this paper, the structure and ideas of Che are explained in more detail, the second part of this pa-
per covers general arguments about this approach and gives a detailed overview of reachable benefits for education.

@@

Openshift\footnote{https://www.openshift.com/} provides a software-suite around Kubernetes with the goal to automate cloud-native development and delivery.
Notable  additions  include  Jenkins,  Gitea,  Sonarqube and lately Che. 
Regarding Che there are no notable changes in handling, as Che is simply run on the build in Kubernetes.
When run on Kubernetes or Docker Che will require two additional containers for authentification with KeyCloack. 
For Openshift the standard-openshift authentification is used.  
Eclipse Che  can be run  on any  base-technology  on  localhost – this is rather for demonstrations aswell as for developers working on the Che-Code. 
When run (locally) on docker, the plugin CheDir enables a portable workspace, 
which can be used on any docker based Che.
\subsection{Technology and Workflow}
\label{subsec:Tec}
A normal IDE is software on your computer which makes programming easier than doing it in a plain text-editor and commandline. 
Common tasks for an IDE are dependency management, debugging, refactoring and autocompletion.
It is important that an IDE is not required to develop – you can write .java-files in Vim, compile on the console and run a .Jar manually. 
With that in mind, we can see that the IDE is just a tool to make things (a lot) more comfortable – but you also have requirements such as the JDK, the .Jars you need in your classpath etc. as they are only utilized by your IDE.
You can consider this stack as your workspace as it contains everything required to work, with the IDE making the workspace comfortable to use.
Eclipse Che does not work like a normal IDE. 
Che consists of three main components shown in Figure \ref{fig:cheworkflow}\footnote{source: \url{https://en.wikipedia.org/wiki/Eclipse_Che#/media/File:Eclipse_Che_-_Workflow.PNG}}.

@@

The university of Arizona showed \cite{CO16}, that about 50\% of software inspected was not able to build. 

@@

Experienced programmers are usually able to overcome these issues, given enough time. This continues to be an issue \cite{CL14}, and without a different approach probably always will be. 
This problem is solved, if the software is built inside a docker-container and follows the principles of self-containment and image-immutability \cite{BI17}.

@@

Che therefore embraces the core-ideas of DevOps \cite{HUE12}\cite{JA16} and agile software development \cite{EBE16}.

@@

This is one of the primary goals of the DevOps approach \cite{JA16}. 
The workspace can therefore form totally new ways of programming, such as IoT Applications \footnote{he Plugin Arktik for Eclipse offered IoT Development with Che.While Arktik is unsupported since February 2019, it proved that the concept worked.} or a full setup of microservices \cite{IV18}, while being agile and reproducible.

@@

Clark et al. \cite{CL14} describe this rather as a change in the mindset, than in actual technology:

@@

Another problem with Che is the uncommon debugging. 
As the code is running in a remote-runtime, the common breakpoint  and  valuecheck
is  missing.  
Instead,  one  either must be familiar with the languages commandline-debugging or write good logs and metrics. 
While this will slow down development, it also offers some benefits: 
The debugging and error-analysis inside the container is the same as it would be in the productive environment. 
Having meaningful logs with different loglevels and collecting valuable metrics is 
in general a good attribute for software in any circumstance. 
It just means additional effort and know-how.  
The last problem is the vulnerability to malevolent users. 
Someone who already is in possession of a valid account can drain the shared resources of a full cluster, therefore blocking  any  other  user  from  work.
This  can  also  be achieved by human error. 
Dependent on the setup, the attacker could also auto-deploy hidden spyware into a pro-
duction environment. 
While this is also a problem for normal developer teams, the fast pace of deployments with Che requires additional care. 
A rich CI/CD environment, such as a well-maintained Jenkins may lead to less-detailed code-reviews, as both unit- and end-to-end-tests succeed.  
It is therefore important to not only restrict access to project resources, but also the double-check and guard the releases.  
Some security risks can also emerge from docker-images as well \cite{PE13} but are out of scope for this paper.

@@

Even complicated topics, like microservices and IoT-applications can easily be adaptet inside Che \cite{IV18} and performed as for example as group-projects with a common goal. 

@@

With both the product and the code inside the workspace easily accessible, the professor can grant faster and better feedback for the students, which can have massive impact on learning \cite{HU95}\cite{HU18}.

@@

This immediate feedback is a big motivator and has a positive impact on the learning experience \cite{FI05}. 
With a clear reason on the usefulness of the task, the visible progress and the direct impact of personal effort should lead to intrinsic motivation, which studies show to have a strong effect on education \cite{DECI93}. 

@@

Other studies have shown that only automated and unified testing is not beneficial to education in general \cite{TC88}. 

@@

The topic of socio-economic factors is beyond the scope of this work, but the impact has been analyzed by the PISA Studies \cite{PISA15}, which state that socio-economicly handicapped students fail three times more likely to achieve a basic level of performance.

@@

Germany scored on the impact of socioeconomic-factors on education better than it did in 2006 but is worse than global average \cite{OECDCN15}.

@@

Making  assignments  mandatory, providing extrinsic motivation \cite{DECI93} has usually a negative impact on the subjective experience.

@@

Long waiting times disconnect the action from the result being not beneficial for the learning process \cite{HU95}\cite{DECI93}. 

@@

With studies showing basics of good learning experiences \cite{FI05}\cite{TC88} and the impact of motivation \cite{HU95}\cite{DECI93}, it is mandatory for institutions to provide the best possible environment. 
